Data-driven approaches to identifying knowledge gaps in marine invertebrate biodiversity

### 2017-04-04 manuscript.Rmd
edit manuscript

With increasing impact of human activities on the biosphere, we need efficient and accurate approaches to document changes in species ranges and community composition. While terrestrial ecosystems have received considerable attention, the baseline of healthy and resilient marine ecosystems are not well understood. A major hinderance is that we do not have enough information about which species live where, and there is potentially a large portion of marine biodiversity that remains to be discovered and described. Additionally, marine ecosystem monitoring are lacking to detect, and even less understand, how and why communities change through time. There is therefore a pressing need to improve our knowledge of marine systems before species go extinct, and their functionning are irreversibly altered. Our lack of knowledge about marine invertebrates make the modelisation of trophic networks less accurate, as diverse species assemblages are often reduced as simple functional categories that do no account for their ecological differences. Until very recently, our knowledge of biodiversity was scattered among heterogeneous sources (e.g. natural history collections, publications, expert knowledge), and difficult to access. This heterogeneity and the difficulty to accessing this data have made synthesis and the identification of knowledge gaps difficult. However, with the development of linked, publicly accessible, comprehensive databases of natural history records (e.g. iDigBio), species occurrences (e.g., GBIF, OBIS), genetic data (e.g., GenBank, BOLD), and registers of species names (e.g., WoRMS), it is now easier to quantitavely assess the state of our knowledge for biodiversity data. Here, we argue that given the urgency of the biodiversity crisis, we need to develop a data-driven approach to speed up the documentation of biodiversity and build on our existing knowledge to be more efficient at collecting information to improve our understanding of our natural world. Additionally, we argue that our understanding biodiversity in the marine environment and the impact of human activities on communities requires not only to get a more accurate list of the species that inhabits the oceans, but also to understand better their distributions. This can only be done with additional sampling and a commitment to training more taxonomists and experts that can identify the samples collected. To some extant, we are at a time when it has never been easier and cost effective to survey comprehensively the extant of marine biodiversity.

Biodiversity discovery and documentation has been transformed in the 21st century by using molecular data to facilitate species identification, delineate species limits, and more recently to quantify diversity and uncover large scale patterns of biodiversity organization. The wide adoption of DNA barcoding (where homologous DNA sequences are sequenced across the main domains of the Eukaryotes) has unraveled the pervasiveness of cryptic species, invigorating taxonomy and the description of new species since the early 2000s. Unraveling these cryptic species has also contributed to the recognition that species have smaller ranges than their taxonomic identity had implied making them potentially more susceptible to extinction or overexploitation. And despite early criticisms, DNA barcoding has also been suggested as a possible solution to the widening gap between the increased need to document biodiversity while taxonomic expertise is dwindling. Thus, large databases of DNA barcodes have been collaboratively assembled and made pubicly available. By matching DNA barcodes of unidentified species to a reference public database of accurately identified and vouchered specimens, DNA barcoding can greatly speed up species identification. Sequences with no match make it easier to flag specimens that warrant further taxonomic attention.

Barcoding/metabarcoding certainly not perfect methods, but much better than current understanding.

@@

Recent developments in high-throughput sequencing technologies are giving new applications to these DNA barcode databases. Metabarcoding has been developed to sequence in parallel the barcoding gene across the species included in environmental samples such as plankton tows, fish guts, or artificial structure scrappings. With decreasing sequencing costs, and the prospect of automating sample collection, these technologies will soon be used for ecosystem and biodiversity monitoring. DNA barcoding provides the possibility to link these newly generated data with historical natural history knowledge about species and their distributions. However, the insights we will gain from these approaches are only going to be as good as the level of completeness and accuracy of the information included in reference databases. However, to accurately identify the species found in these samples, comprehensive DNA barcode databases are needed. Here, we focus mostly on museum specimens because we can double check their identification and we can potentially extract DNA from them to complete existing databases of DNA barcodes.

@@

Our knowledge about species and their distributions is gathered in regional faunistic lists. They are typically assembled by taxonomic experts. They are comprehensive but usually limited in their geographic scope. Natural history collections have more comprehensive geographic coverage, but the quality of the taxonomic identifications can vary.

Here, we assess the level of completeness of biodiversity public databases for marine invertebrates that occur within the boundaries of the EEZ of the US. Specifically, we use comprehensive regional faunistic lists to assess how much of this knowledge is captured by current digitization efforts of museum collections and occurrence records; and we estimate how many of the species have genetic data publicly available that would be useful in the context of identifying samples from metabarcoding studies.

>>>>

With increasing impact of human activities on the biosphere, we need efficient and accurate approaches to document changes in species ranges and community composition. While terrestrial ecosystems have received considerable attention, the baseline of healthy and resilient marine ecosystems are not well understood, and modeling the impact of human activities is limited to a few taxonomic groups. A major hinderance is that we do not have enough information about which species live where, and an unknown but potentially significant portion of marine biodiversity remains to be discovered and described. Additionally, marine ecosystem monitoring initiatives are lacking to detect and understand, how and why communities change through time [@Duffy2013,@Borja2016]. There is therefore a pressing need to improve our knowledge of marine systems before species go extinct, and their functionning are irreversibly altered. Our lack of knowledge about marine invertebrates make the modelisation of trophic networks less accurate, as diverse species assemblages are often reduced as simple functional categories that do no account for their ecological diversity. Until very recently, our knowledge of biodiversity was scattered among heterogeneous sources (e.g. natural history collections, publications, expert knowledge), and difficult to access. This heterogeneity and the difficulty to access this data have made synthesis and the identification of knowledge gaps difficult. However, with the development of linked, public, comprehensive databases of natural history records (e.g. iDigBio), species occurrences (e.g., GBIF, OBIS), genetic data (e.g., GenBank, BOLD), and registers of species names (e.g., WoRMS), it is now easier to quantitavely evaluate the state of our knowledge for biodiversity data. Here, we aggregate publicly available data for marine invertebrates that occur in the United States to highligth gaps in our knowledge. Given the urgency of the biodiversity crisis, we need to develop data-driven initiatives to efficiently document biodiversity, and build on our existing knowledge to collect information to improve our understanding of marine systems. Additionally, we emphasize that our understanding of marine biodiversity, and the impact of human activities on communities require not only to get a more accurate list of the species that inhabits the oceans, but we also need better data on species distributions and abundances. This can only be achieved with additional sampling and a commitment to training more taxonomists and experts that can identify the material collected. We live at a time when it has never been easier and cost effective to survey comprehensively the extant of marine biodiversity.

Biodiversity discovery and documentation has been transformed in the 21st century by using molecular data to facilitate species identification, delineate species. And more recently, the development of high throughput sequencing technologies is used to quantify diversity and uncover large scale patterns of biodiversity organization. The wide adoption of DNA barcoding (where homologous DNA sequences are sequenced across the main domains of the Eukaryotes) has unraveled the pervasiveness of cryptic species, invigorating taxonomy and the description of new species, and enabled collaborative research programs such as the Census of Marine Life. Unraveling these cryptic species has also contributed to the recognition that species have smaller ranges than their taxonomic identity had implied making them potentially more susceptible to extinction or overexploitation. Despite early criticisms and now well-understood limitations, DNA barcoding has been suggested as a possible solution to the widening gap between the increased need to document biodiversity while taxonomic expertise is dwindling. Thus, large databases of DNA barcodes have been collaboratively assembled and made pubicly available. By matching DNA barcodes of unidentified species to a reference public database of accurately identified and vouchered specimens, DNA barcoding can greatly speed up species identification. Sequences with no match make it easier to flag specimens that warrant further taxonomic attention.

@@

Recent developments in high-throughput sequencing technologies are giving new applications to these DNA barcode databases [@Leray2016]. Metabarcoding has been developed to sequence in parallel the barcoding gene across the species included in environmental samples such as plankton tows, fish guts, or artificial structure scrappings [@Leray2015]. With decreasing sequencing costs, and the prospect of automating sample collection, these technologies will soon be used routinely for ecosystem and biodiversity monitoring [@Bohan2017]. DNA barcoding provides the possibility to link these newly generated data with historical natural history knowledge about species and their distributions. However, the insights we will gain from these approaches are only going to be as good as the level of completeness and accuracy of the information included in reference databases. However, to accurately identify the species found in these samples, comprehensive DNA barcode databases are needed. Here, we mostly focus on museum specimens because we can validate their identification, and a fraction of these samples (if they have not been preserved in formalin) can be used to obtain DNA from them to complete existing databases of DNA barcodes.

<!-- Barcoding/metabarcoding certainly not perfect methods, but much better than current understanding. -->

@@

Our knowledge about species and their distributions is gathered in regional faunistic lists, typically assembled by taxonomic experts. They are comprehensive but limited in their geographic scope. On the other hand, natural history collections have more comprehensive geographic coverage, but the quality of the taxonomic identifications can vary, and some common species might be underrepresented. We assessed the level of completeness of biodiversity public databases for marine invertebrates that occur within the boundaries of the EEZ of the US. Specifically, we compared comprehensive regional faunistic lists to assess how much of this knowledge is captured by current digitization efforts of museum collections and occurrence records; and we estimate how many of the species have genetic data publicly available that would be useful in the context of identifying samples from metabarcoding studies.
